\documentclass{beamer}
%\usepackage{ngerman}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
\usepackage{lmodern}
\usepackage{pgf}
\usepackage{verbatim}
\usepackage{listings}
\usepackage[]{hyperref} 
\usepackage{multicol}
\usepackage{wasysym}
\usepackage{tikz}

\setbeamersize{text margin left = 0.4em}
\setbeamersize{text margin right = 0.4em}
%\setbeamercovered{transparent}

\title{Restful interface for Kolab}
\subtitle{bachelor thesis}
\institute{Fernuniversität Hagen}
\date{\today}
\author{Thomas Koch}

\usetheme{boxes}
\usecolortheme{dolphin}

\AtBeginSection[]
{
   \begin{frame}
       \frametitle{Outline}
       \tableofcontents[currentsection]
   \end{frame}
}

\begin{document}
\begin{frame}
  \maketitle{}
\end{frame}

% \begin{frame}{me}
%   \begin{itemize}
%   \item 5 years dev: PHP, Java, Hadoop, Crawler, CMS
%   \item Relations: Debian, FSFE, Basic Income, ATTAC
%   \item around 2006: eGroupWare
%   \item Full time student again
%   \end{itemize}
% \end{frame}

\begin{frame}{Restful PIM}
  \begin{itemize}
  \item Life is too short for WebDAV (CardDAV, CalDAV)
  \item Design and prototype a restful groupware API
  \end{itemize}
\end{frame}

\begin{frame}{Do you need\ldots}
  \begin{itemize}
  \item locking (remember CVS?)
  \item versioning (for PIM data?)
  \item additional HTTP verbs?
  \item TooMuchSpecException, incompatibilites?
  \item Push notifications? WebDAV hasn't. Atom has.
  \end{itemize}
\end{frame}

\begin{frame}{What's restful?}
  \ldots more then just HTTP!

  Rule of thumb: If your API needs documentation, it's not restful.

  \begin{itemize}
  \item Just one entrance URI
  \item Every concept (resource) identified by it's own discovarable URI
  \item Just four methods: GET, PUT, DELETE, POST
  \item All the magic in well known Media Types and Link Relations
    \begin{itemize}
    \item Atom feed (collection), xCal, HTML page, HAL - Hypertext Application Language
    \item author, hub, search, replies, next, prev, up, copyright
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Rob Yates \url{robubu.com}: CalATOM, CardATOM}
\begin{itemize}
  \item existing standards: HTTP, Atom, AtomPub
  \item widely deployed / implemented
  \item kind of known by many developers
  \item many optional features: Push notification, OpenID, oauth
  \item principles not limited to PIM data: Project Management, Photo gallery
\end{itemize}
\end{frame}

\begin{frame}
\begin{quote}
Were I to propose CalDAV today it would probably be CalAtom  
\end{quote}
-- Lisa Dusseault, February 29, 2008
\end{frame}

\begin{frame}{Synchronize}
  \begin{quote}
    Consume all entries of a (paged\footnote{RFC 5005 - Feed Paging and
      Archiving}) feed unless you see an entry older then the newest entry from
    your last synchronization run.
  \end{quote}
  Delete with tombstones (The Atom "deleted-entry" Element, draft-snell-atompub-tombstones)
\end{frame}

\begin{frame}{Edit while online}
  \begin{itemize}
  \item Check resource for update / retrieve etag
  \item User edits
  \item Conditional PUT request
  \item Synchronize
  \end{itemize}
\end{frame}

\begin{frame}{Edit while offline}
  \begin{itemize}
  \item Save changed resource localy, keep original
  \item Go online
  \item Synchronize
  \item No conflict: Conditional PUT request for each changed resource
  \item For Conflicts: Present 3-way merge to user (3 windows?)
  \end{itemize}
\end{frame}

\begin{frame}{Why restful HTTP instead of IMAP?}
  \begin{itemize}
  \item HTTP (also) has security, parallel downloads, persistent connections, caching, compression, download continuation via ranges
  \item Lisa Dusseault: HTTP Access to Email Stores \url{http://tools.ietf.org/id/draft-dusseault-httpmail-00.txt}
  \item google://rest advantages (loose coupling, evolvable, caching, simple)
  \item Read Fielding's thesis! (Chapters 4-5)
  \end{itemize}
\end{frame}
\end{document}
